Photosynthesis Course Note

### 2016-07-24 Draft/draft.tex
Adding notes for A.Zee's QFT in a Nutshell

\usepackage{float}  % For H in figures
\usepackage{amsmath, amssymb} % For math
\usepackage{mathtools} % dcases*, see https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics#The_cases_environment
\numberwithin{equation}{subsection} % have the enumeration go to the subsection level.
									% See:https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics
\usepackage{graphicx}   % need for figures
\usepackage{cite} % For bibligraphy
\usepackage{fancyref} % For lazy reference \fref
\usepackage{hyperref} % For hyperlink everything.
\usepackage{CJKutf8} % For Chinese characters
%\usepackage{ dsfont } % For double struck fonts
\usepackage{braket} 
\usepackage[T1]{fontenc}
\usepackage{listings}

@@

>>>>

\usepackage{float}  % For H in figures
\usepackage{amsmath, amssymb} % For math
\usepackage{mathtools} % dcases*, see https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics#The_cases_environment
\numberwithin{equation}{subsection} % have the enumeration go to the subsection level.
									% See:https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics
\usepackage{graphicx}   % need for figures
\usepackage{cite} % For bibligraphy
\usepackage{fancyref} % For lazy reference \fref
\usepackage[unicode]{hyperref} % For hyperlink everything.
\usepackage{CJKutf8} % For Chinese characters
%\usepackage{ dsfont } % For double struck fonts
\usepackage{braket} 
\usepackage[T1]{fontenc}
\usepackage{listings}

@@

    \subsubsection{2.4 The Poincar\'{e} Algebra}
    \label{sec:2.4_The_Poincare_Algebra}
    \textbf{I am lost in this part.} The general idea seems to develop, in a 
    infinitesimal sense, the property of a Lorentz transformation.
    For a infinitesimal Lorentz transformation:
    \begin{align}
        U(1+\omega, \epsilon) = 
            1+ \frac{1}{2} i\omega_{\rho\sigma}J^{\rho\sigma}
            -i\epsilon_\rho P^\rho + \cdots
    \end{align}
    where $J$ and $P$ are independent of the infinitesimal value
    $\omega$ and $\epsilon$.

    Later $P$ is identified as the energy-momentum operator (with
    $P^0$ being the energy operator), and $J^{23}$, $J^{31}$,
    \(J^{12}\) are identified with the angular momentum operator.
    However, the reason for this identification is not provided. 
    % TODO why P is energy momentum, and J is angular momentum

    Later, it was eatablished that $J$ and $P$ satisfy:
    \begin{align}
        i[J^{\mu\nu},J^{\rho\sigma}] &=
            \eta^{\nu\rho}J^{\mu\sigma} - \eta^{\mu\rho}J^{\nu\sigma}
            -\eta^{\sigma\mu}J^{\rho\nu} + \eta^{\sigma\nu}J^{\rho\mu}
        \\
        i[P^\mu,J^{\rho\sigma} &=
            \eta^{\mu\rho}P^\sigma - \eta^{\mu\sigma}P^{\rho}
        \\
        [P^\mu,P^\rho] &= 0
    \end{align}
    This is called the \textit{Lie Algebra} of the Poincar\'{e}
    group.
    The relation \([P^\mu,P^\rho] = 0\) is particular interesting.

    With this, the finite translation and a rotation by angle
    $\theta$ are expressed as:
    \begin{align}
        U(1,a) = \exp(-iP^\mu a_\mu) \\
        U(\theta,0) = \exp(i \mathbf{J}\cdot \theta)
    \end{align}

    This part ends with a discussion on the low-velocity limit
    of the Lie Algebra obtained above, i.e. the Galilean algebra.

    \subsection{Note}
    \label{sec:Note}
    I will now change to another book: Q.F.T. in a Nutshell by A. Zee.

\section{Quantum Field Theory in a Nutshell}
\label{sec:Quantum_Field_Theory_in_a_Nutshell}
    \subsection{Part I: Motivition and Foundations}
    \label{sec:QFT_in_a_Nutshell.Part_I}
        \subsubsection{I:2 Path Integral Formulation of Quantum Physics}
        \label{sec:QFT_in_a_Nutshell_Part_I.2}
        Here the path integral formulation of quantum mechanics is introduced.
        The intuition comes from a limiting case of the traditional double 
        slit electron interference experiment. (\textbf{pp.7 to 10}) Then 
        it calculates the transition probability $\braket{q_F|e^{-iHT}|q_I}$
        by divide it into a infinite of steps:
        \begin{align}
            \braket{q_F|e^{-iHT}|q_I} = \lim_{N\to \infty}
            \braket{q_F|e^{iH\delta t}e^{iH\delta t}\cdots e^{iH\delta t}|q_I}
            (\text{with }N\delta t = T)
        \end{align}

        For illustration, it calculates this value when 
        $H=\frac{\hat{p}^2}{2m}$. The result is that:
        \begin{align*}
            \braket{q_F|e^{-iHT}|q_I} = \int Dq(t) 
                e^{i\int_0^T dt \frac{1}{2}m \dot{q}^2 }
        \end{align*}

        where:
        \begin{align}
            \int Dq(t) \equiv \lim_{N\to \infty} 
            \left( \frac{-im}{2\pi \delta t}\right)^{N/2}
            \left( \prod_{k=1}^{N-1} \int dq_k \right)
        \end{align}
        It notes that when $H=\hat{p}^2/2m + V(\hat{q})$, the final result
        would have been:
        \begin{align}
            \label{eq:QFT_in_a_Nutshell_Part_I.2.TransProbability}
            \braket{q_F|e^{-iHT}|q_I} &= \int Dq(t) 
                e^{i\int_0^T dt \frac{1}{2}m \dot{q}^2 - V(q)}
                \nonumber \\
                &= e^{i \int_0^T dt L(\dot{q}, q)}
        \end{align}
        where $L$ is the Lagrangian of the system.

        The value \(\braket{0|e^{-iHT}|0}\) is denoted $Z$. This part
        mentions that one often effect a change of coordinate
        $t \to -it$, called \textit{Wick rotation}, to obtain:
        \begin{align}
            Z = \int Dq(t) e^{-\int_0^T dt H(\dot{q},q)}
        \end{align}
        where $H$ is the Hamiltonian of the system.
        The mathematical rigorious aspect is often ignored.

        It also discuss how this formulation could explain the classical
        limit of quantum mechanics, i.e. classical mechanics, in a
        very direct manner. This is related to the saddle point
        approximation to the integral 
        \ref{eq:QFT_in_a_Nutshell_Part_I.2.TransProbability}.

        \paragraph{Unclear point}
        Why is $\int dq \ket{q}\bra{q} = 1$ while $\int \frac{p}{2\pi}
        \ket{p}\bra{p} = 1$. What does it mean by saying
        "to see that the normalization is correct" (pp. 10 and 11).

        \paragraph{Corresponding pages in draft} pp. 1 to 3.

\section{Miscellnaneous Math}
\label{sec:Miscellnaneous_Math}
\subsection{\texorpdfstring{$\int_0^\infty \frac{\sin(x)}{x}$}{}}
From \href{http://math.stackexchange.com/questions/5248/solving-the-integral-int-0-infty-frac-sinxx-dx-frac-pi2}{Math.SE}. By Aryabhata (http://math.stackexchange.com/users/1102/aryabhata).

I believe this can also be solved using double integrals.

It is possible (if I remember correctly) to justify switching the order of integration to give the equality:

$$\int_{0}^{\infty} \Bigg(\int_{0}^{\infty} e^{-xy} \sin x \,dy \Bigg)\, dx = \int_{0}^{\infty} \Bigg(\int_{0}^{\infty} e^{-xy} \sin x \,dx \Bigg)\,dy$$
Notice that
$$\int_{0}^{\infty} e^{-xy} \sin x\,dy = \frac{\sin x}{x}$$

This leads us to

$$\int_{0}^{\infty} \Big(\frac{\sin x}{x} \Big) \,dx = \int_{0}^{\infty} \Bigg(\int_{0}^{\infty} e^{-xy} \sin x \,dx \Bigg)\,dy$$
Now the right hand side can be found easily, using integration by parts.

\begin{align*}
I &= \int e^{-xy} \sin x \,dx = -e^{-xy}{\cos x} - y \int e^{-xy} \cos x \, dx\\
&= -e^{-xy}{\cos x} - y \Big(e^{-xy}\sin x + y \int e^{-xy} \sin x \,dx \Big)\\
&= \frac{-ye^{-xy}\sin x - e^{-xy}\cos x}{1+y^2}.
\end{align*}
Thus $$\int_{0}^{\infty} e^{-xy} \sin x \,dx = \frac{1}{1+y^2}$$
Thus $$\int_{0}^{\infty} \Big(\frac{\sin x}{x} \Big) \,dx = \int_{0}^{\infty}\frac{1}{1+y^2}\,dy = \frac{\pi}{2}.$$

### 2016-07-24 Light harvesting/Photosynthesis Course Note.tex
Adding notes for A.Zee's QFT in a Nutshell

\section{July 6th Wed}
\label{sec:July_6th}
This lecture begins with a note on the unbelievable efficiency of biological
system. Then it proceeds to review some concepts in quantum statistical 
mechanics ( \textbf{page 2 to 21}). Since we are quite familar with this
subject, it will be omitted from this note. The last page mentioned an
important equation for the evolvement of density matrix in an open system,
called the "master equation".
% TODO reference required!

@@

Next, we explore several interesting properties of these coherent
states:
\begin{itemize}
    \item \textbf{Expectations}
        (\textbf{pp.16 to 17}) It is found that 
        $\braket{X}=\sqrt{2}\lambda \text{Re}(\alpha)$,
        and $\braket{P} = \frac{\sqrt{2}\hbar}{\lambda} \text{Im}(\alpha)$. Or:
        $\alpha$ is related to $\braket{\alpha|X|\alpha}$
        and $\braket{\alpha|P|\alpha}$. Hence a real $\alpha$ means the state
        $\ket{\alpha}$ is a static state with $\braket{P}=0$. Moreover, this 
        state
        is the lowest energy eigenstate of a simple harmonic oscillator with
        the origin shifted from $0$ to $\alpha$. 
        % todo substantiate this point!
    \item \textbf{Variance}
        (\textbf{pp.18 to 20})
        It is also found that the variance 
        $(\Delta X)^2 = \frac{\lambda^2}{2}$,
        this is exactly the same as the variance of the gound state of simple
        harmonic oscillator.
        And $(\Delta P)^2 = \frac{\hbar^2}{2\lambda^2}$. 
        From the two expression,
        we see that any coherent states have the same minimal uncertainty:
        $\Delta X \cdot \Delta P = \frac{\hbar}{2}$.
    \item \textbf{Time Evolution}
        (\textbf{pp.21})
        Using the energy eigenstate of simple harmonic oscillator, we can find 
        the time evolution of coherent state as 
        $\alpha(t) = e^{-i\omega t} \alpha_0$. It shows that $\alpha$ is really
        like a \textit{classical oscillator} and the norm $|\alpha|^2$ is conserved. 
        This is so cool!
\end{itemize}

@@

Next we write out the Hamiltonian for the LH1/2. (\textbf{pp.26})
The phonon $H_{ph}$ term 
contains 16 phonons for 16 pigments in LH1. The exciton term $H_{ex}$
assumes that the interaction is only nearst neighbour interaction.
The exciton-phonon interaction $H_{\text{ex-ph}}$ is essentially
the usual electron-phonon interaction term in solid state physics.
\section{July 15, Friday}
\label{sec:July_15}

@@

\textbf{pp.8} introduces the Davydov's ansatz.
Note that
\begin{align*}
    e^{A+B} = e^A e^B e^{\frac{1}{2}[A,B]}
\end{align*}
% todo check this
And using this relationship, it could be found that the $\ket{\beta}$
is actually a Glauber coherent state.

@@

Explain for the energy band.
% todo level repulsion

\textbf{pp.10 }shows a Holstein polaron's phase diagram. 
\footnote{Could be found in pdf file "nonlocal.pdf".}
The region enclosed by the pointed solid line is of particular
interests. Outside this thin region,
there is always a unique solution (global minimum for the variational
parameters). 
% Below the region, large polaron. Above, small polaron.
% todo what is a large polaron.
Outside the region, there will be two solutions coexisting.
The the solution obatined depeands on the initial configuration.
The physical picture is that, changing the value of $J$ and $g$,
the size of polaron will change sharply when one transverse the
dashed line.

@@

The remaining slides re not be included in this note,
since they follow the pattern of applying models developed
by condensed matter physicsists and showing the result.
Detailed results could be found in the papers provided by
the teacher.
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.

>>>>

\section{July 6th Wed}
\label{sec:July_6th}
This lecture begins with a note on the unbelievable efficiency of biological
system. Then it proceeds to review some concepts in quantum statistical 
mechanics ( \textbf{page 2 to 21}). Since we are quite familar with this
subject, it will be omitted from this note. The last page mentioned an
important equation for the evolvement of density matrix in an open system,
called the "master equation".\footnote{Suggested reading: 
PhysRevA.62.033821}

@@

Next, we explore several interesting properties of these coherent
states:
\begin{itemize}
    \item \textbf{Expectations}
        (\textbf{pp.16 to 17}) It is found that 
        $\braket{X}=\sqrt{2}\lambda \text{Re}(\alpha)$,
        and $\braket{P} = \frac{\sqrt{2}\hbar}{\lambda} \text{Im}(\alpha)$. Or:
        $\alpha$ is related to $\braket{\alpha|X|\alpha}$
        and $\braket{\alpha|P|\alpha}$. Hence a real $\alpha$ means the state
        $\ket{\alpha}$ is a static state with $\braket{P}=0$. Moreover, this 
        state
        is the lowest energy eigenstate of a simple harmonic oscillator with
        the origin shifted from $0$ to $\alpha$. 
        % TODO substantiate this point!
    \item \textbf{Variance}
        (\textbf{pp.18 to 20})
        It is also found that the variance 
        $(\Delta X)^2 = \frac{\lambda^2}{2}$,
        this is exactly the same as the variance of the gound state of simple
        harmonic oscillator.
        And $(\Delta P)^2 = \frac{\hbar^2}{2\lambda^2}$. 
        From the two expression,
        we see that any coherent states have the same minimal uncertainty:
        $\Delta X \cdot \Delta P = \frac{\hbar}{2}$.
    \item \textbf{Time Evolution}
        (\textbf{pp.21})
        Using the energy eigenstate of simple harmonic oscillator, we can find 
        the time evolution of coherent state as 
        $\alpha(t) = e^{-i\omega t} \alpha_0$. It shows that $\alpha$ is really
        like a \textit{classical oscillator} and the norm $|\alpha|^2$ is conserved. 
        This is so cool!
\end{itemize}

@@

Next we write out the Hamiltonian for the LH1/2. (\textbf{pp.26})
The phonon $H_{ph}$ term 
contains 16 phonons for 16 pigments in LH1. The exciton term $H_{ex}$
assumes that the interaction is only nearst neighbour interaction.
The exciton-phonon interaction $H_{\text{ex-ph}}$ is essentially
the usual electron-phonon interaction term in solid state physics.
\section{July 15th, Friday}
\label{sec:July_15}

@@

\textbf{pp.8} introduces the Davydov's ansatz.
Note that
\begin{align*}
    e^{A+B} = e^A e^B e^{\frac{1}{2}[A,B]}
\end{align*}
% TODO check this
And using this relationship, it could be found that the $\ket{\beta}$
is actually a Glauber coherent state.

@@

\textbf{pp.10 }shows a Holstein polaron's phase diagram. 
\footnote{Could be found in pdf file "nonlocal.pdf".}
The region enclosed by the pointed solid line is of particular
interests. Outside this thin region,
there is always a unique solution (global minimum for the variational
parameters). 
% Below the region, large polaron. Above, small polaron.
% TODO what is a large polaron.
Outside the region, there will be two solutions coexisting.
The the solution obatined depeands on the initial configuration.
The physical picture is that, changing the value of $J$ and $g$,
the size of polaron will change sharply when one transverse the
dashed line.

@@

The remaining slides are not included in this note,
because I felt sleepy at that moment. Thankfully the
attached papers have all the interesting information.
\section{July 20th, Wed}
\label{sec:July_20th_Wed}
In this lecture, the author first explains that the $q=0$ mode is
de-coupled into a simple Hamiltonian. Then he uses the time-dependent
variation method developed by Dirac, Frenkel and McLachlan. This method
solves the equation:
\begin{align}
    \label{eq:}
    \braket{\delta \Psi| i\hbar \dot{\Psi} - H\Psi} = 0
\end{align}
with the constraint of $\Psi$ being normalized.

It is also established (by McLachlan) that minimize the following
value:
\begin{align}
    \label{eq:July_20th_Wed}
    ||i\dot{\Psi}-H\Psi||
\end{align}
is an alternative option.

Next we introduces three different Ansatzes: $D_1$ (\textbf{pp.3})
$D_2$ and $\tilda{D}$ (\textbf{pp.4}). The $D_1$ ansatz is most
numerical demanding. It accouts for the difference of phonons in
different site of the 
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
